\tool: Revisiting BadUSB with Type-C

### 2021-03-19 WOOT/Files/background.tex

%\noindent\outline{HID Protocol}\\
Right after one year of the appearance of \ac{USB} 1.0, a standard named \acf{HID}~\cite{hid} was designed based on \ac{USB}. \ac{HID} is
designed with to unify the implementation for devices like mice,
keyboards, etc. Before its appearance, the standard is divided between
manufacturers, for example, the mouse of Company A may use X-Y coordinates to
represent its location while the mouse of Company B uses relative displacement.
This means every device needs its own driver to work. After \ac{HID}, users only need to
write one driver for an entire class of \acp{HID}. Furthermore, \ac{HID} standard also
requires all devices to be \ac{PnP}, which is indeed
convenient but insecure too.

@@

%\noindent\outline{USB2.0}\\
In 2000, \ac{USB} 2.0~\cite{usb20} specification was released. With high speed \mbox{(480
Mbit/s)} mode introduced, printers, cameras, CD-ROM drives, and network cards were
supported in this revision. Such a high data transfer rate also gave rise to the
popularity of ``flash drive'', a portable device that allows physically
transferring data around~\cite{sok}. Although various peripherals were supported
in \ac{USB} 2.0, there was no reliable way to identify the type of device. This
security flaw allowed attacks like BadUSB~\cite{badusb,rubber}.

@@

%\noindent\outline{Security Problem}\\
During the development of \ac{USB} specification, security was insufficiently considered~\cite{sok}. 
The \ac{USB}-IF believes it is the duty of \acp{OEM}
to decide whether security features should be implemented~\cite{usbsec}. But the divergent implementations give a chance for attacks like
BadUSB~\cite{rubber} and our \tool.

>>>>

%\noindent\outline{HID Protocol}\\
Right after one year of the appearance of \ac{USB} 1.0, a standard named \acf{HID}~\cite{hid} was designed based on \ac{USB}. \ac{HID} is
designed with to unify the implementation for devices like mice,
keyboards, etc. Before its appearance, the standard is divided between
manufacturers, for example, the mouse of Company A may use X-Y coordinates to
represent its location while the mouse of Company B uses relative displacement.
This means every device needs its own driver to work. After \ac{HID}, users only need to
write one driver for an entire class of \acp{HID}. Furthermore, the \ac{HID} standard also
requires all devices to be \ac{PnP}, which is indeed
convenient but insecure too.

@@

%\noindent\outline{USB2.0}\\
In 2000, the \ac{USB} 2.0~\cite{usb20} specification was released. With high speed \mbox{(480
Mbit/s)} mode introduced, printers, cameras, CD-ROM drives, and network cards were
supported in this revision. Such a high data transfer rate also gave rise to the
popularity of ``flash drive'', a portable device that allows physically
transferring data around~\cite{sok}. Although various peripherals were supported
in \ac{USB} 2.0, there was no reliable way to identify the type of device. This
security flaw allowed attacks like BadUSB~\cite{badusb,rubber}.

@@

%\noindent\outline{Security Problem}\\
During the development of the \ac{USB} specification, security was insufficiently considered~\cite{sok}. 
The \ac{USB}-IF believes it is the duty of \acp{OEM}
to decide whether security features should be implemented~\cite{usbsec}. But the divergent implementations give a chance for attacks like
BadUSB~\cite{rubber} and our \tool.

### 2021-03-19 WOOT/Files/related_work.tex

To solve this flaw, GoodUSB~\cite{tian2015defending} reports the functionality claimed
by the device to the user and let the user decide whether to authorize. When a device
is plugged in, GoodUSB blocks its functionality before enumeration until a
series of authorizations are completed. These authorizations are designed to be
performed manually, thus the malicious devices will be detected if they make a different claim 
that does not match the authorization results. 
USBeSafe\cite{usbesafe} analyses the 
feature vector constructed with the data collected from USB Request Blocks, 
detects the novel \ac{USB} packets based on the machine learning algorithm and sends 
an alert to the user before enumeration. As BadUSB attacks normally request privileged 
interfaces during enumerating, these defenses are sufficient for these attacks.
Moreover, Mueller \emph{et al.}~\cite{MuellerZN19} improve the security without changing user experience. It can detect the presence of the user, block malicious devices in the user's absence until the session is unlocked.

After the \ac{USB} enumeration and driver loading, some related works follow the \textit{defense in depth} approach to defend against BadUSB attacks.  
Neuner \emph{et al.}~\cite{neuner2018usblock}
prevent BadUSB attacks from the malicious flash drive by analyzing
the temporal characteristics of BadUSB-like attacks. This defense mechanism is
effective because the attacker cannot obtain the screen of the host using
BadUSB. In this case, the malicious device can only inject keystrokes in a very
short time to reduce the risk of being discovered. This defect causes the
typing characteristics of BadUSB to be detectable. Pham \emph{et al.}~\cite{pham2010optimizing} optimize Windows security features, which can
block the execution of unsigned files and the installation of unsigned drivers
carried on portable media. Moreover, in GoodUSB, a VM is deployed in the host as
a honeypot to detect and stop malicious behavior of \ac{USB} devices.

In addition to injection attacks, data theft attack is also one of the focuses
of the academic community. As mentioned in Section~\ref{subsec:usb_attack}, there
exists an attack called JFA~\cite{JFC} which abuses the \ac{MHL}
standard to steal video stream from the victim. In order to mitigate this
issue, Meng et al.~\cite{meng2018252} propose a statistical model using status
like GPU/CPU usage to detect JFA attacks.

>>>>

To solve this flaw, GoodUSB~\cite{tian2015defending} reports the functionality claimed
by the device to the user and lets the user decide whether to authorize. When a device
is plugged in, GoodUSB blocks its functionality before enumeration until a
series of authorizations are completed. These authorizations are designed to be
performed manually, thus the malicious devices will be detected if they make a different claim 
that does not match the authorization results. 
USBeSafe\cite{usbesafe} analyses the 
feature vector constructed with the data collected from USB Request Blocks, 
detects the novel \ac{USB} packets based on the machine learning algorithm and sends 
an alert to the user before enumeration. As BadUSB attacks normally request privileged 
interfaces during enumerating, these defenses are sufficient for these attacks.
Moreover, Mueller \emph{et al.}~\cite{MuellerZN19} improve the security without changing user experience. It can detect the presence of the user, block malicious devices in the user's absence until the session is unlocked.

After the \ac{USB} enumeration and driver loading, some related works follow the \textit{defense in depth} approach to defend against BadUSB attacks.  
Neuner \emph{et al.}~\cite{neuner2018usblock}
prevent BadUSB attacks from the malicious flash drive by analyzing
the temporal characteristics of BadUSB-like attacks. This defense mechanism is
effective because the attacker cannot obtain the screen of the host using
BadUSB. In this case, the malicious device can only inject keystrokes in a very
short time to reduce the risk of being discovered. 
This property causes the
typing characteristics of BadUSB to be detectable. Pham \emph{et al.}~\cite{pham2010optimizing} optimize Windows security features, which can
block the execution of unsigned files and the installation of unsigned drivers
carried on portable media. Moreover, in GoodUSB, a VM is deployed in the host as
a honeypot to detect and stop malicious behavior of \ac{USB} devices.

In addition to injection attacks, data theft attack is also one of the focuses
of the academic community. As mentioned in Section~\ref{subsec:usb_attack}, there
exists an attack called JFA~\cite{JFC} which abuses the \ac{MHL}
standard to steal the video stream from the victim. In order to mitigate this
issue, Meng et al.~\cite{meng2018252} propose a statistical model using status
like GPU/CPU usage to detect JFA attacks.
